---
title: "DATA 606 Data Project Proposal"
author: "Beshkia Kvarnstrom"
date: '`r Sys.Date()`'
output:
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

### LOAD PACKAGES
```{r, message=FALSE, warnings=FALSE}
pkges <- c("tidyverse", "readr", "kableExtra", "dplyr", "ggplot2", "lubridate", "psych")

# Loop through the packages
for (p in pkges) {
  # Check if package is installed
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p) #If the package is not installed, install the package
    
    library(p, character.only = TRUE) #Load the package
  } else {
    library(p, character.only = TRUE) #If the package is already installed, load the package
  }
}
```

### Load Data from GitHub
```{r, message=FALSE, warnings=FALSE}
TikTok_Data <- read.csv("https://raw.githubusercontent.com/BeshkiaKvarnstrom/MSDS-DATA606/main/sug_users_vids_all.csv")

glimpse(TikTok_Data)
```

### Subset the columns in the DataSet
```{r, message=FALSE, warnings=FALSE}

TikTokDF <- TikTok_Data[,c("user_name", "create_time", "hashtags", "video_length", "n_likes", "n_shares", "n_comments", "n_plays", "n_followers", "n_total_likes", "n_total_vids")]

#Use the Rename() function to change the column names in the dataframe 
TikTokDF <- TikTokDF %>% 
           rename("TikTok_User_Name" = "user_name", "Video_Date" = "create_time", "Hashtags" = "hashtags", "Video_Length" = "video_length", "Total_Video_Likes" = "n_likes", "Total_Video_Shares" = "n_shares", "Total_Video_Comments" = "n_comments", "Total_Video_Plays"= "n_plays", "Total_User_Followers" = "n_followers", "Total_User_Likes" = "n_total_likes", "Total_Videos" = "n_total_vids")

#Add new Total Engagement column, which is the sum of all likes, shares, comments, and views
TikTokDF <- TikTokDF %>%
  group_by(TikTok_User_Name) %>% #Here we group the major category
mutate(Total_Engagement = Total_Video_Likes+Total_Video_Shares+Total_Video_Comments+Total_Video_Plays, Engagement_Rate = (Total_Video_Likes+Total_Video_Shares+Total_Video_Comments+Total_Video_Plays)/Total_User_Followers, Video_Date = as_datetime(Video_Date) )

# Remove duplicate rows
TikTokDF <- TikTokDF %>% distinct()
TikTokDF

```

## Research Question 
TikTok, owned by a Chinese company called ByteDance, has grown exponentially and has gained popularity as one of the most popular social media app (with over 700 million users) that allows users to create, watch, and share short videos on any topic. The length of TikTok videos ranges in duration from 3 seconds to 10 minutes. The platform allows users to get creative with their content using filters, stickers, voice overs, sound effects, and background music.

I choose TikTok for my discussion because of its popularity as well as its rapid growth over the past 2 years and its use of video marketing marketing to pique its users interest almost to the point of addiction to the platform.. I am also intrigued by TikTok's “For You” feed where each user's experience is tailored to your own interests.

Hence my research question is: Does the duration of a TikTok video influences it popularity positively or negatively?

## What are the cases, and how many are there? 
The dataset contains data on engagement with user-created videos and user profile data including the video length, likes, share and comments which will be analyzed in this project. There are 41682 observations/cases.
```{r, message=FALSE, warnings=FALSE}
nrow(TikTokDF)
```

## Describe the method of data collection.
I found the TikTok dataset on Kaggle. It was originally obtained from TikTok's trending API by a GitHub user named Ivan Tran.

## What type of study is this (observational/experiment)?
I will be performing an observational study of the dataset.


## Data Source: If you collected the data, state self-collected. If not, provide a citation/link.
I found the TikTok dataset on Kaggle. The dataset is linked here: https://www.kaggle.com/datasets/vbradculbertson/tiktok-trending-metadata?select=sug_users_vids_all.csv


## Response: What is the response variable, and what type is it (numerical/categorical)?
The response variable is 'Engagement' and is numerical. Engagement is the sum of all likes, shares, comments, and views.


## Explanatory: What is the explanatory variable(s), and what type is it (numerical/categorival)?
The explanatory variables are Video Length(in seconds) and Total Videos(Number of videos uploaded by the creator).


## Relevant summary statistics 
```{r, message=FALSE, warnings=FALSE}
#Total_Engagement independent variable 
summary(TikTokDF$Total_Video_Plays)
```

```{r, message=FALSE, warnings=FALSE}
#Total_Engagement independent variable 
summary(TikTokDF$Total_Engagement)
```

```{r, message=FALSE, warnings=FALSE}
describeBy(TikTokDF$Total_Engagement, group = TikTokDF$TikTok_User_Name, mat = TRUE)
```
```{r, message=FALSE, warnings=FALSE}
#The following scatter plot depicts the relationship between The Total TikTok Videos and Comments 
ggplot(TikTokDF, aes(Total_Videos, Total_Video_Comments)) +
  geom_point(size=1, shape=23,  col="purple") + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE, color = 'green') +
labs(x="Total Videos", y="Total Comments", 
     title="Relationship between The Total TikTok Videos and Comments")


```

```{r, message=FALSE, warnings=FALSE}
#The following scatter plot depicts the relationship between The Length of TikTok Videos and Comments 
ggplot(TikTokDF, aes(Video_Length, Total_Video_Comments)) +
  geom_point(size=1, shape=23,  col="purple") + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE, color = 'green') +
labs(x="Video Length(in seconds)", y="Total Comments", 
     title="Relationship between The length of TikTok Videos and Comments")


```

